Exploring dimensions of biodiversity in Japanese ferns

### 2020-06-30 ms/SI.Rmd
Update MS and SI
add modified t-test to biplots

```{r plot-prep}
tax_regions_data <- regions_taxonomy %>%
  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

phy_regions_data <- regions_phylogeny %>%
  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

ses_div_ferns_spatial <-
  shape_ferns %>%
  left_join(categorize_endemism(ses_phy_ferns), by = c(grids = "site")) %>%
  left_join(ses_traits_ferns, by = c(grids = "site")) %>%
  mutate(redundancy = 1 - (richness/abundance)) %>%
  left_join(tax_regions_data, by = "grids") %>%
  left_join(phy_regions_data, by = "grids")

ses_div_ferns_endemic_spatial <-
shape_ferns %>%
  left_join(categorize_endemism(ses_phy_ferns_endemic), by = c(grids = "site"))

map_theme <- theme(
    panel.grid.major = element_line(color = "white", size = 0.2),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
  )
```

@@

>>>>

```{r plot-prep}
# Combine bioregion results, biodiversity metrics, and shape data into single dataframe
tax_regions_data <- regions_taxonomy %>%
  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  select(grids, taxonomic_cluster)

phy_regions_data <- regions_phylogeny %>%
  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  select(grids, phylo_cluster)

ses_div_ferns_spatial <-
  shape_ferns %>%
  left_join(ses_phy_ferns, by = c(grids = "site")) %>%
  left_join(ses_traits_ferns, by = c(grids = "site")) %>%
  left_join(tax_regions_data, by = "grids") %>%
  left_join(phy_regions_data, by = "grids")

# - same for endemic species
ses_div_ferns_endemic_spatial <-
shape_ferns %>%
  left_join(ses_phy_ferns_endemic, by = c(grids = "site"))

# Define ggplot2 theme for maps
map_theme <- theme(
    panel.grid.major = element_line(color = "white", size = 0.2),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
  )

@@

```{r biplot-prep}
# Extract centroids
centroids <- sf::st_centroid(ses_div_ferns_spatial$geometry) %>%
  sf::st_coordinates()

# Define helper function for making a scatterplot
scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate) {
  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.5) +
    labs(
      x = x_lab,
      y = y_lab
    )

  # Add annotation and trendline if annotate = TRUE
  if(annotate) p <- p +
      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, fill = "transparent") +
      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)

  p
}
```

```{r raw-biplot, fig.width = 7, fig.height = 6}
## Biplots of raw metrics ##

# Make dataframe of arguments to feed into plotting function
# - includes all non-duplicated combinations of raw diversity metrics
raw_comparisons <-
  cross_df(
    list(
      x_var = c("mpd_obs", "pe_obs", "mpd_morph_obs", "richness"),
      y_var = c("mpd_obs", "pe_obs", "mpd_morph_obs")
    ),
    .filter = `==` # filter out exact duplicates
  ) %>%
  # Remove reversed duplicates
  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
  # Use factors to keep the preferred combination of x and y
  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs", "mpd_morph_obs", "richness")) %>%
  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs", "pe_obs", "mpd_morph_obs")) %>%
  arrange(x_var, y_var) %>%
  slice(1) %>%
  ungroup() %>%
  select(-grp) %>%
  # Convert x_var and y_var back to character for plotting function
  mutate(across(c(x_var, y_var), as.character)) %>%
  # Add labels
  mutate(
    x_lab = case_when(
      x_var == "mpd_morph_obs" ~ "MMD",
      x_var == "mpd_obs" ~ "MPD",
      x_var == "pe_obs" ~ "PE",
      x_var == "richness" ~ "Richness"
    ),
    y_lab = case_when(
      y_var == "mpd_morph_obs" ~ "MMD",
      y_var == "mpd_obs" ~ "MPD",
      y_var == "pe_obs" ~ "PE",
      y_var == "richness" ~ "Richness"
    )
  ) %>%
  # Run modified t-test
  mutate(
    test_res = map2(
      .x = x_var, 
      .y = y_var, 
      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
    # Extract relevant stats from results
    cor = map_dbl(test_res, pluck("corr")),
    pval = map_dbl(test_res, pluck("p.value"))
    ) %>%
  # Generate label
  mutate(
    cor_lab = glue("r = {round(cor, 3)}"),
    pval_lab = glue("p = {round(pval, 3)}"),
    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
  ) %>%
  # Specify which plots to annotate
  mutate(
    annotate = case_when(
      x_var == "mpd_morph_obs" & y_var == "mpd_obs" & pval < 0.05 ~ TRUE,
      x_var == "mpd_morph_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
      x_var == "mpd_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
      x_var == "richness" & y_var == "pe_obs" & pval < 0.05  ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  select(x_var, y_var, x_lab, y_lab, label, annotate)

# Make plot
pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
wrap_plots() +
  plot_annotation(tag_levels = 'A')

ggsave(result_file("raw-biplot", "pdf"), width = 7, height = 6, units = "in")
```

**`r s_figure("raw-biplot")`**. `r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism. Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05; correlations between MPD and MMD with richness not expected to be linear, so were not tested.

```{r ses-biplot, fig.width = 7, fig.height = 6}
## Biplots of SES metrics ##

# Make dataframe of arguments to feed into plotting function
# - includes all non-duplicated combinations of raw diversity metrics
ses_comparisons <-
  cross_df(
    list(
      x_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z", "richness"),
      y_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")
    ),
    .filter = `==` # filter out exact duplicates
  ) %>%
  # Remove reversed duplicates
  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
  # Use factors to keep the preferred combination of x and y
  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs_z", "mpd_morph_obs_z", "richness")) %>%
  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")) %>%
  arrange(x_var, y_var) %>%
  slice(1) %>%
  ungroup() %>%
  select(-grp) %>%
  # Convert x_var and y_var back to character for plotting function
  mutate(across(c(x_var, y_var), as.character)) %>%
  # Add labels
  mutate(
    x_lab = case_when(
      x_var == "mpd_morph_obs_z" ~ "SES MMD",
      x_var == "mpd_obs_z" ~ "SES MPD",
      x_var == "pe_obs_z" ~ "SES PE",
      x_var == "richness" ~ "Richness"
    ),
    y_lab = case_when(
      y_var == "mpd_morph_obs_z" ~ "SES MMD",
      y_var == "mpd_obs_z" ~ "SES MPD",
      y_var == "pe_obs_z" ~ "SES PE",
      y_var == "richness" ~ "Richness"
    )
  ) %>%
  # Run modified t-test
  mutate(
    test_res = map2(
      .x = x_var, 
      .y = y_var, 
      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
    # Extract relevant stats from results
    cor = map_dbl(test_res, pluck("corr")),
    pval = map_dbl(test_res, pluck("p.value"))
    ) %>%
  # Generate label
  mutate(
    cor_lab = glue("r = {round(cor, 3)}"),
    pval_lab = glue("p = {round(pval, 3)}"),
    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
  ) %>%
  # Specify which plots to annotate
  mutate(
    annotate = case_when(
      pval < 0.05 ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  select(x_var, y_var, x_lab, y_lab, label, annotate)

# Make plot
pmap(ses_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
  wrap_plots() +
  plot_annotation(tag_levels = 'A')

ggsave(result_file("ses-biplot", "pdf"), width = 7, height = 6, units = "in")
```

**`r s_figure("ses-biplot")`**. `r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism. Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; all combinations tested, but *p* value and correlation (r) only shown for *p* < 0.05.

### 2020-06-30 ms/manuscript.Rmd
Update MS and SI
add modified t-test to biplots

>>>>

### Relationships between metrics

Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`D, E). Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`F). Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`A, `r s_figure("ses-biplot")`A, B). SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`A, B, C). None of the SES values were correlated with richness (`r s_figure("ses-biplot")`D, E, F).
